A Poll-of-Polls Forecast for the 2024 U.S. Presidential Election: Analyzing Support Trends for Harris by Pollster and State

### 2024-11-04 paper/paper.qmd
refinement of 'measurement' and 'references'

Each poll represents a predictor of an actual event, namely voter opinion at a particular moment. Nevertheless, like all survey results, the raw data is susceptible to many potential limitations including the following: sampling error, variation in polling methods, distortion because of inappropriate survey responses such as missing data or response from respondents who misunderstood the questions [@tellingstories].

While applying several filters to the original dataset such as restricting to those with a numeric grade of 2.7 or higher or pollsters with more than 30 polls improves data reliability, certain limitations still exist. Selection bias and sampling error remains as a concern, since polls always represent only part of the population. Differences in the way different organizations conducted their polling might introduce more inconsistencies. Finally, by focusing our attention on post-declaration polls only, we exclude earlier trends that could add more insight into how Harris' support has evolved over time.

@@

Our modelling strategy estimates Kamala Harris's support percentage in the 2024 US election polls, accounting for potential variations over time including pollster and state differences. The model balances complexity with interpretability, using both linear and Bayesian frameworks to capture patterns in the data. We turned from the initial linear model to a Bayesian model for robust predictions considering the variables' uncertainties. The Bayesian approach, which incorporates both prior beliefs and observed data to make predictions, allowed us to capture the variances among pollsters and states more accurately and responsively. Further background details and diagnostics are included in [Appendix -@sec-model-details].

@@

These models are run through the rstanarm package [@rstanarm] in R  [@citeR], which makes Bayesian modeling available through the use of Stan's strong inference engine. To validate the models, RMSE and WAIC have been made use of to check the goodness of fit; Bayesian models with reduced RMSE and WAIC outperform linear models. We use weakly informative priors; for example, $\beta_0 \sim \text{Normal}(0, 10)$ and $\sigma \sim \text{Exponential}(1)$. This reflects our initial uncertainty but prevents overfitting. The priors were chosen conservatively to ensure that the model remains consistent.

Model diagnostics, including posterior predictive checks and convergence diagnostics, were carried out to ensure the reliability of the results. The Bayesian models converged successfully, as indicated by $\hat{R} = 1$ for all parameters.

The main assumption in these models is that the pollster and state effects can be treated as random. This assumes that the effects are normally distributed across pollsters and states, which may not always be accurate. Additionally, the model assumes that polling data is representative of the actual electorate, an assumption that can be violated if polls are biased or have non-random sampling issues. Despite these limitations, the hierarchical structure allows us to capture important variability, making the model suitable for predicting Harris's support. Future improvements could involve incorporating time-varying effects or exploring interactions between pollsters and states.

@@

\

@@

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults1
#| tbl-cap: "Linear models of support percentages for Harris based on date and pollster"
#| warning: false
#| tbl-width: 80% 
#| page-break-after: true
#| page-break-before: true

@@

The proposed survey questionnarie design is in the following link:
\
https://forms.gle/zQ8iJyPk3HhJYrKd9.

::: {.layout}
### Survey demo

::: {.column width="1/5"}
![Survey intro](../other/idealized_survey/survey_intro.png)
:::

::: {.column width="1/5"}
![Survey questions asking about demographics](../other/idealized_survey/survey_q1.png)
:::

::: {.column width="1/5"}
![Survey questions asking for voting intentions and candidate favorability](../other/idealized_survey/survey_q2.png)
:::

::: {.column width="1/5"}
![Survey questions asking for issues of interest and likelihood to vote](../other/idealized_survey/survey_q3.png)
:::

::: {.column width="1/5"}
![Survey final thanks](../other/idealized_survey/survey_outro.png)
:::
:::

>>>>

In converting an individual's complex set of opinions into "plausibly measurable" [@tellingstories] numbers, each poll attempts to capture and simplify voters' choice in a straightforward way. Polls capture a momentary snapshot of opinion, but this process is inevitably influenced by factors such as question phrasing, the timing of the poll, and contextual events. A person may have mixed or conflicting feelings about candidates, and the process of capturing these preferences in a single response imposes much error [@tellingstories].

The outcome variable, which is intended to reflect overall support levels, is constructed from respondents' stated candidate preferences. As part of the data refinement process to enhance the reliability of predictions, we restricted our analysis to polls with a numeric quality grade of 2.7 or higher and to pollsters with over 30 polls in the dataset. We set a minimum threshold of 2.7 for the numeric grade to focus on the top 30% of the dataset, ensuring a sufficient sample size of 5,612 observations. Additionally, we included only pollsters with at least 30 polls, reasoning that a higher number of polls from a given pollster would enhance the reliability and validity of the aggregated data.

While efforts to filter and refine the dataset mitigate some limitations, they cannot entirely eliminate the challenges inherent in measuring multifaceted concept like voter preference. Limitations including sampling error, response bias, and inconsistencies in polling methods remain. Some responses may also be distorted due to misunderstandings or incomplete answers, as respondents might interpret questions differently depending on context [@tellingstories]. This measurement process is essentially an estimation, translating complex human sentiments into quantifiable data, subject to various limitations.

@@

Our modelling strategy estimates Kamala Harris's support percentage in the 2024 US election polls, accounting for potential variations over time including pollster and state differences. The model balances complexity with interpretability, using both linear and Bayesian frameworks to capture patterns in the data. We turned from the initial linear model to a Bayesian model for robust predictions considering the variables' uncertainties. The Bayesian approach, which incorporates both prior beliefs and observed data to make predictions, allowed us to capture the variances among pollsters and states more accurately and responsively. Further background details and diagnostics are included in @sec-model-details.

@@

These models are run through the rstanarm package [@rstanarm] in R [@citeR]. RMSE and WAIC were used to check for the goodness of fit of the model; Bayesian models with reduced RMSE and WAIC outperform linear models. We use weakly informative priors; for example, $\beta_0 \sim \text{Normal}(0, 10)$ and $\sigma \sim \text{Exponential}(1)$. This reflects our initial uncertainty but prevents overfitting. The priors were chosen conservatively to ensure that the model remains consistent.

Model diagnostics, including posterior predictive checks and convergence diagnostics, were carried out to ensure the reliability of the results. The Bayesian models converged successfully, as indicated by $\hat{R} = 1$ for all parameters. Further details about model diagnostics are included in @sec-model-details.

The main assumption in these models is that the pollster and state effects can be treated as random. This assumes that the effects are normally distributed across pollsters and states, which may not always be accurate. Additionally, the model assumes that polling data is representative of the actual electorate, an assumption that can be violated if polls are biased or have non-random sampling issues. Despite these limitations, the hierarchical structure allows us to capture important variability, making the model suitable for predicting Harris's support. Future improvements could involve incorporating time-varying effects or interaction terms between pollsters and states.

@@

@@

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults1
#| tbl-cap: "Linear models of support percentages for Harris based on date and pollster"
#| warning: false
#| tbl-width: 60% 
#| tbl-height: 55%
#| page-break-after: true
#| page-break-before: true

@@

The proposed survey questionnarie design is in the following link:
\
https://forms.gle/vehDvpDNDGZ8WCn47

::: {.container}
### Survey demo

::: {.column width="30%"}
![Survey Intro](../other/idealized_survey/survey_intro.png)
:::

::: {.column width="30%"}
![Questions asking for vote eligibility - citizenship, age](../other/idealized_survey/eligibility_1.png)
:::

::: {.column width="30%"}
![Questions asking for vote eligibility - voter status and registration state](../other/idealized_survey/eligibility_2.png)
:::

::: {.column width="100%"}
::: {.columns}
::: {.column width="33%"}
![Demographics Questions](../other/idealized_survey/survey_q1.png)
:::
::: {.column width="33%"}
![Voting Intentions and Candidate Favorability](../other/idealized_survey/survey_q2.png)
:::
::: {.column width="33%"}
![Issues of Interest and Likelihood to Vote](../other/idealized_survey/survey_q3.png)
:::
:::
:::

::: {.column width="30%"}
![Survey Final Thanks](../other/idealized_survey/survey_outro.png)
:::

:::
